\textbf{Selective DNA and Protein Isolation from Marine Macrophyte Surfaces}

### 2020-11-19 submission/manuscript.tex
Format customization.

@@

\sisetup{mode=text}
\setlength\parindent{24pt}

@@

To evaluate whether the obtained DNA is suitable to determine the
composition of the microbial community Illumina sequencing of the V4
region of the 16S rRNA was performed. Sequencing yielded a total of
336,944 sequences after quality curation and exclusion of eukaryotic,
mitochondrial and no relative sequences. The number of sequences
classified as chloroplasts was 97,334. After excluding these sequences
the total number of retrieved reads was 239,610, ranging from 22,596 to
52,930 sequences per sample (\autoref{supp-nseq_notus}). Even when the
highest sequencing effort was applied, the rarefaction curves did not
level off which is commonly observed in high-throughput 16S rRNA
amplicon sequencing (\autoref{supp-rarefaction}). Sequences clustering
at a similarity level of 97 \si{\percent} yielded a total of 8,355
different OTUs. Taxonomic classification of reads revealed a
macrophyte-associated epiphytic community that mainly composed of
\emph{Alphaproteobacteria} (14.9 ± 3.5 \si{\percent}),
\emph{Bacteroidota} (12.5 ± 2.4 \si{\percent}),
\emph{Gammaproteobacteria} (11.6 ± 4.3 \si{\percent}),
\emph{Desulfobacterota} (7.8 ± 7.5 \si{\percent}), \emph{Cyanobacteria}
(6.5 ± 4.7 \si{\percent}) and \emph{Planctomycetota} (2.9 ± 1.7
\si{\percent}) (\autoref{community}).

Primers used in this study, as in many other 16S rRNA amplicon
procedures, also amplified chloroplast 16S rRNA genes. We observed a
high proportion of chloroplast sequences in all analyzed samples (33.4 ±
9.4 \si{\percent}) (\autoref{community}). To determine whether
chloroplast sequences originate from the host or eukaryotic epiphytic
organisms, we exported SILVA-classified chloroplast sequences and
reclassified them using the RDP (Ribosomal Database Project) training
set that allows for a more detailed chloroplast classification. The
largest proportion of sequences was classified as Bacillariophyta (89.7
± 5.7 \si{\percent}) indicating that the DNA removal procedure resulted
in only minor co-extracted quantities of host DNA
(\autoref{chloroplast}). Chloroplast sequences classified as
Streptophyta constituted 3.3 ± 2.8 \si{\percent} of all chloroplast
sequences originating from \emph{C. nodosa} samples, while sequences
classified as Chlorophyta comprised only 0.02 ± 0.01 \si{\percent} of
all chloroplast sequences associated with \emph{C. cylindracea} samples.

To determine whether the extracted DNA can be used for metagenomic
sequencing, four samples containing epiphytic DNA were selected and
shotgun sequenced using an Illumina platform. Metagenomic sequencing
yielded between 207,149,524 and 624,029,930 sequence pairs
(\autoref{supp-metagenomic_statistics}). Obtained sequences were
successfully assembled into contigs whose L50 ranged from 654 to 1,011
bp. In addition, predicted coding sequences were functionally annotated
(9,066,667 -- 20,256,215 annotated sequences; \autoref{cog}) and
taxonomically classified. Functional annotation allowed for an
assessment of the relative contribution of each COG (Clusters of
Orthologous Groups) functional category to the total number of annotated
coding sequences (\autoref{cog}a). Functional categories containing the
highest number of sequences were C (Energy production and conversion), E
(Amino acid transport and metabolism), M (Cell wall/membrane/envelope
biogenesis), L (Replication, recombination and repair) and P (Inorganic
ion transport and metabolism). If host DNA is co-extracted with
epiphytes it should be detected in large proportions in sequenced
metagenomes. However, no large proportions of coding sequences
classified as Streptophyta and Chlorophyta were detected
(\autoref{supp-metagenomic_taxonomy}). Sequenced metagenomic DNA
originating from the surface of \emph{C. nodosa} contained 1.3
\si{\percent} of coding sequences classified as Streptophyta in December
2017 and 0.7 \si{\percent} in June 2018. Furthermore, the summed RPKM
(Reads Per Kilobase Million) of these sequences constituted 1.7
\si{\percent} of total RPKM of all successfully classified sequences in
December 2017 and 1.1 \si{\percent} in June 2018. Similar low
proportions of host coding sequences were detected in metagenomic
samples originating from the surfaces of \emph{C. cylindracea}. Of all
successfully classified coding sequences 0.2 \si{\percent} were
classified as Chlorophyta in December 2017 and 0.1 \si{\percent} in June
2018. A relatively higher proportion of RPKM of these sequences than in
the case of \emph{C. nodosa} was observed, indicating a higher
coextraction of host DNA in \emph{C. cylindracea}. In December, the
proportion of RPKM of sequences classified as Chlorophyta was 8.2
\si{\percent}, while in June 2018 it reached 13.6 \si{\percent}.

@@

To test whether the developed DNA and protein isolation protocols
efficiently detach microbes from the macrophyte surface we selected
\emph{C. nodosa} and \emph{C. cylindracea} as representatives of
seagrass and macroalgal species. These species differ morphologically.
While \emph{C. nodosa} leaves are flat, \emph{C. cylindracea} thalli are
characterized by an uneven surface (Kuo and den Hartog, 2001; Verlaque
\emph{et al.}, 2003). The developed protocol led to an almost complete
removal of epiphytic cells from the surfaces of both species comparable
to the result of Burke \emph{et al.} (2009), indicating that structural
differences do not impact the removal efficiency. In addition, isolation
protocols were tested in two contrasting seasons, as it is known that
macrophytes are harboring more algal epiphytes during autumn and winter
(Reyes and Sansón, 2001). No differences in the removal efficiency was
observed between seasons suggesting that these protocols can be used on
macrophyte samples retrieved throughout the year. Also, no removal
differences were observed on samples derived from the same host but from
different locations.

@@

The DNA was isolated according to the protocol for isolation from
filters described in Massana \emph{et al.} (1997). This protocol was
modified and adapted for microbial DNA isolation from macrophyte
surfaces as described below. Five (5) \si{\ml} of lysis buffer (40
\si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar}
sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or 2
\si{\g} wet-weight of thalli. For every sample, duplicate isolations
were performed. Lysozyme was added (final concentration 1
\si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius}
for 30 \si{\minute}. Subsequently, proteinase K (final concentration 0.5
\si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were
added and the samples were incubated at 55 \si{\degreeCelsius} for 2
\si{\hour}. Following the incubation, tubes were vortexed for 10
\si{\minute} and the mixture containing lysed epiphytic cells was
separated from host leaves or thalli by transferring the solution into a
clean tube. The lysate was extracted twice with a mixture of
phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with
chloroform:isoamyl alcohol (24:1). After each addition of an organic
solvent mixture, tubes were slightly vortexed and centrifuged at 4,500 ×
g for 10 \si{\minute}. Following each centrifugation the aqueous phases
were retrieved. After the final extraction 1/10 of chilled 3 \si{\Molar}
sodium acetate (pH 5.2) was added. DNA was precipitated by adding 1
volume of chilled isopropanol, incubating the mixtures overnight at
\num{-20} \si{\degreeCelsius} and centrifuging at 16,000 × g and 4
\si{\degreeCelsius} for 20 \si{\minute}. The pellet was washed twice
with 1 \si{\ml} of chilled 70 \si{\percent} ethanol and centrifuged
after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 10
\si{\minute}. After the first washing step duplicate pellets from the
same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.
The dried pellet was resuspended in 100 \si{\ul} of deionized water.

@@

An aliquot of isolated DNA was treated with RNase A (final concentration
200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
concentration was determined using the Quant-iT PicoGreen dsDNA Assay
Kit (Invitrogen, USA) according to the manufacturer's instructions and
diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
amplified using a two-step PCR procedure. In the first PCR the 515F
(5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3')
primers from the Earth Microbiome Project
(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
were used to amplify the target region (Caporaso \emph{et al.}, 2012;
Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
contained on their 5' end a tagged sequence. Each sample was amplified
in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5
Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml}
BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse
primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs,
USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial
denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of
denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
volumes were pooled and PCR products were purified using the GeneJET PCR
Purification Kit (Thermo Fisher Scientific, USA) according to the
manufacturer's instructions and following the protocol that included
isopropanol addition for better small DNA fragment yield. The column was
eluted in 30 \si{\ul} of deionized water. Purified PCR products were
sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
Martinsried, Germany. Before sequencing at IMGM, the second PCR
amplification of the two-step PCR procedure was performed using primers
targeting the tagged region incorporated in the first PCR. In addition,
these primers contained adapter and sample-specific index sequences. The
second PCR was carried out for 8 cycles. Beside samples, a positive and
negative control were sequenced. A negative control was comprised of
four parallel PCR reactions without DNA template, while for a positive
control a mock community composed of evenly mixed DNA material
originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
used. Partial 16S rRNA sequences obtained in this study have been
deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under
accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794,
SAMEA6648809 -- SAMEA6648811.

Obtained sequences were analyzed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommendations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
SILVA were exported and reclassified using mothur and the RDP
(\url{http://rdp.cme.msu.edu/}) training set (version 16) reference
files adapted for mothur (Cole \emph{et al.}, 2014). In comparison to
SILVA, RDP allows a more detailed classification of chloroplast
sequences. Based on the ATCC MSA-1002 mock community included in the
analysis a sequencing error rate of 0.009 \si{\percent} was determined,
which is in line with previously reported values for next-generation
sequencing data (Kozich \emph{et al.}, 2013; Schloss \emph{et al.},
2016). In addition, the negative control processed together with the
samples yielded only 2 sequences after sequence quality curation.

@@

Obtained sequences were analyzed on the Life Science Compute Cluster
(LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
Individual sequences were assembled using MEGAHIT (version 1.1.2) (Li
\emph{et al.}, 2015) under default settings. Putative genes were
predicted from contings longer than 200 bp using Prodigal (version
2.6.3) (Hyatt \emph{et al.}, 2010) in metagenome mode (-p meta).
Abundances of predicted genes were expressed as Reads Per Kilobase
Million (RPKM) and calculated using the BWA algorithm (version 0.7.16a)
(Li and Durbin, 2009). All predicted genes were functionally annotated
using the eggNOG-mapper (Huerta-Cepas \emph{et al.}, 2017) and eggNOG
database (version 5.0) (Huerta-Cepas \emph{et al.}, 2019). Sequence
taxonomy classification was determined using the lowest common ancestor
algorithm adapted from DIAMOND (version 0.8.36) (Buchfink \emph{et al.},
2015) and by searching against the NCBI non-redundant database (NR). To
determine the phylogeny, the top 10 \si{\percent} hits with an e-value
\textless{} 1 × 10\textsuperscript{-5} were used (-\/-top 10). Sequence
renaming, coverage information computing and metagenomic statistics
calculations were performed using software tools from BBTools
(\url{https://jgi.doe.gov/data-and-tools/bbtools}). Metagenomic
sequences obtained in this study have been deposited in the European
Nucleotide Archive (ENA) at EMBL-EBI under accession numbers
SAMEA6648795, SAMEA6648797, SAMEA6648809 and SAMEA6648811.

@@

Isolated proteins were whole trypsin digested using the FASP
(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
according to the manufacturer's instructions with small modifications
(Wiśniewski \emph{et al.}, 2009). Prior to loading the solution onto the
column, protein pellets were solubilized in a urea sample buffer
included in the kit amended with DTT (final concentration 100
\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
remove larger particles. The first washing step after protein solution
loading was repeated twice. In addition, the centrifugation steps were
prolonged if the column was clogged. Trypsin digestion was performed on
column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
The final filtrate containing peptides was acidified with 1
\si{\percent} (final concentration) trifluoroacetic acid (TFA), freezed
at \num{-80} \si{\degreeCelsius}, lyophilized and sent to VIME -- Vienna
Metabolomics Center (University of Vienna) for metaproteomic analysis.
Peptides were resuspended in 1 \si{\percent} (final concentration) TFA,
desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA)
according to the manufacturer's instructions and sequenced on a Q
Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher
Scientific, USA). Obtained MS/MS spectra were searched against a protein
database composed of combined sequenced metagenomes using SEQUEST-HT
engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
Fisher Scientific, USA). The target-decoy approach was used to reduce
the probability of false peptide identification. Results whose false
discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
For protein identification a minimum of two peptides and one unique
peptide were required. For protein quantification, a chromatographic
peak area-based free quantitative method was applied.

@@

Host leaves and thalli from DNA and protein isolation steps were washed
seven times in deionized water and fixed with formaldehyde (final
concentration \textasciitilde{} 3 \si{\percent}). In addition,
nontreated leaves and thalli, washed three times in ASW to remove
loosely attached microbial cells, were fixed in the same concentration
of formaldehyde and used as a positive control. For-long term storage,
fixed leaves and thalli were immersed in a mixture of phosphate-buffered
saline (PBS) and ethanol (1:1) and stored at \num{-20}
\si{\degreeCelsius}. Treated and untreated segments of leaves and thalli
were stained in a 2 × solution of SYBR Green I and examined under a
Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).

@@

\leavevmode\hypertarget{ref-Li2009a}{}%
Li, H. and Durbin, R. (2009) Fast and accurate short read alignment with
BurrowsWheeler transform. \emph{Bioinformatics} \textbf{25}: 1754--1760.

@@

\leavevmode\hypertarget{ref-Verlaque2003}{}%
Verlaque, M., Durand, C., Huisman, J.M., Boudouresque, C.-F., and Le
Parco, Y. (2003) On the identity and origin of the Mediterranean
invasive \emph{Caulerpa} \emph{Racemosa} (Caulerpales, Chlorophyta).
\emph{Eur J Phycol} \textbf{38}: 325--339.

@@

\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\textit{C. nodosa} and \textit{C. cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-3}
\end{figure}

>>>>

\usepackage{setspace}

@@

\newpage
\linenumbers
\sisetup{mode=text}
\setlength\parindent{24pt}
\doublespacing

@@

To evaluate whether the obtained DNA is suitable to determine the
composition of the microbial community Illumina sequencing of the V4
region of the 16S rRNA was performed. Sequencing yielded a total of
336,944 sequences after quality curation and exclusion of eukaryotic,
mitochondrial and no relative sequences. The number of sequences
classified as chloroplasts was 97,334. After excluding these sequences,
the total number of retrieved reads was 239,610, ranging from 22,596 to
52,930 sequences per sample (\autoref{supp-nseq_notus}). Even when the
highest sequencing effort was applied, the rarefaction curves did not
level off which is commonly observed in high-throughput 16S rRNA
amplicon sequencing (\autoref{supp-rarefaction}). Sequences clustering
at a similarity level of 97 \si{\percent} yielded a total of 8,355
different OTUs. Taxonomic classification of reads revealed a
macrophyte-associated epiphytic community that mainly composed of
\emph{Alphaproteobacteria} (14.9 ± 3.5 \si{\percent}),
\emph{Bacteroidota} (12.5 ± 2.4 \si{\percent}),
\emph{Gammaproteobacteria} (11.6 ± 4.3 \si{\percent}),
\emph{Desulfobacterota} (7.8 ± 7.5 \si{\percent}), \emph{Cyanobacteria}
(6.5 ± 4.7 \si{\percent}) and \emph{Planctomycetota} (2.9 ± 1.7
\si{\percent}) (\autoref{community}).

Primers used in this study, as in many other 16S rRNA amplicon
procedures, also amplified chloroplast 16S rRNA genes. We observed a
high proportion of chloroplast sequences in all analysed samples (33.4 ±
9.4 \si{\percent}) (\autoref{community}). To determine whether
chloroplast sequences originate from the host or eukaryotic epiphytic
organisms, we exported SILVA-classified chloroplast sequences and
reclassified them using the RDP (Ribosomal Database Project) training
set that allows for a more detailed chloroplast classification. The
largest proportion of sequences was classified as Bacillariophyta (89.7
± 5.7 \si{\percent}) indicating that the DNA removal procedure resulted
in only minor co-extracted quantities of host DNA
(\autoref{chloroplast}). Chloroplast sequences classified as
Streptophyta constituted 3.3 ± 2.8 \si{\percent} of all chloroplast
sequences originating from \emph{C. nodosa} samples, while sequences
classified as Chlorophyta comprised only 0.02 ± 0.01 \si{\percent} of
all chloroplast sequences associated with \emph{C. cylindracea} samples.

To determine whether the extracted DNA can be used for metagenomic
sequencing, four samples containing epiphytic DNA were selected and
shotgun sequenced using an Illumina platform. Metagenomic sequencing
yielded between 207,149,524 and 624,029,930 sequence pairs
(\autoref{supp-metagenomic_statistics}). Obtained sequences were
successfully assembled into contigs whose L50 ranged from 654 to 1,011
bp. In addition, predicted coding sequences were functionally annotated
(9,066,667 -- 20,256,215 annotated sequences; \autoref{cog}a) and
taxonomically classified. Functional annotation allowed for an
assessment of the relative contribution of each COG (Clusters of
Orthologous Groups) functional category to the total number of annotated
coding sequences (\autoref{cog}a). Functional categories containing the
highest number of sequences were C (Energy production and conversion), E
(Amino acid transport and metabolism), M (Cell wall/membrane/envelope
biogenesis), L (Replication, recombination and repair) and P (Inorganic
ion transport and metabolism). If host DNA is co-extracted with
epiphytes it should be detected in large proportions in sequenced
metagenomes. However, no large proportions of coding sequences
classified as Streptophyta and Chlorophyta were detected
(\autoref{supp-metagenomic_taxonomy}). Sequenced metagenomic DNA
originating from the surface of \emph{C. nodosa} contained 1.3
\si{\percent} of coding sequences classified as Streptophyta in December
2017 and 0.7 \si{\percent} in June 2018. Furthermore, the summed RPKM
(Reads Per Kilobase Million) of these sequences constituted 1.7
\si{\percent} of total RPKM of all successfully classified sequences in
December 2017 and 1.1 \si{\percent} in June 2018. Similar low
proportions of host coding sequences were detected in metagenomic
samples originating from the surfaces of \emph{C. cylindracea}. Of all
successfully classified coding sequences 0.2 \si{\percent} were
classified as Chlorophyta in December 2017 and 0.1 \si{\percent} in June
2018. A relatively higher proportion of RPKM of these sequences than in
the case of \emph{C. nodosa} was observed, indicating a higher
co-extraction of host DNA in \emph{C. cylindracea}. In December, the
proportion of RPKM of sequences classified as Chlorophyta was 8.2
\si{\percent}, while in June 2018 it reached 13.6 \si{\percent}.

@@

To test whether the developed DNA and protein isolation protocols
efficiently detach microbes from the macrophyte surface we selected
\emph{C. nodosa} and \emph{C. cylindracea} as representatives of
seagrass and macroalgal species. These species differ morphologically.
While \emph{C. nodosa} leaves are flat, \emph{C. cylindracea} thalli are
characterized by an uneven surface (Kuo and den Hartog, 2001; Verlaque
\emph{et al.}, 2003). The developed protocol led to an almost complete
removal of epiphytic cells from the surfaces of both species comparable
to the result of Burke \emph{et al.} (2009), indicating that structural
differences do not impact the removal efficiency. In addition, isolation
protocols were tested in two contrasting seasons, as it is known that
macrophytes are harbouring more algal epiphytes during autumn and winter
(Reyes and Sansón, 2001). No differences in the removal efficiency was
observed between seasons suggesting that these protocols can be used on
macrophyte samples retrieved throughout the year. Also, no removal
differences were observed on samples derived from the same host but from
different locations.

@@

The DNA was isolated according to the protocol for isolation from
filters described in Massana \emph{et al.} (1997). This protocol was
modified and adapted for microbial DNA isolation from macrophyte
surfaces as described below. Five (5) \si{\ml} of lysis buffer (40
\si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar}
sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or 2
\si{\g} wet-weight of thalli. For every sample, duplicate isolations
were performed. Lysozyme was added (final concentration 1
\si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius}
for 30 \si{\minute}. Subsequently, proteinase K (final concentration 0.5
\si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were
added and the samples were incubated at 55 \si{\degreeCelsius} for 2
\si{\hour}. Following the incubation, tubes were vortexed for 10
\si{\minute} and the mixture containing lysed epiphytic cells was
separated from host leaves or thalli by transferring the solution into a
clean tube. The lysate was extracted twice with a mixture of
phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with
chloroform:isoamyl alcohol (24:1). After each addition of an organic
solvent mixture, tubes were slightly vortexed and centrifuged at 4,500 ×
g for 10 \si{\minute}. Following each centrifugation the aqueous phases
were retrieved. After the final extraction 1/10 of chilled 3 \si{\Molar}
sodium acetate (pH 5.2) was added. DNA was precipitated by adding 1
volume of chilled isopropanol, incubating the mixtures overnight at
\num{-20} \si{\degreeCelsius} and centrifuging at 16,000 × g and 4
\si{\degreeCelsius} for 20 \si{\minute}. The pellet was washed twice
with 1 \si{\ml} of chilled 70 \si{\percent} ethanol and centrifuged
after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 10
\si{\minute}. After the first washing step duplicate pellets from the
same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.
The dried pellet was re-suspended in 100 \si{\ul} of deionized water.

@@

An aliquot of isolated DNA was treated with RNase A (final concentration
200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
concentration was determined using the Quant-iT PicoGreen dsDNA Assay
Kit (Invitrogen, USA) according to the manufacturer's instructions and
diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
amplified using a two-step PCR procedure. In the first PCR the 515F
(\(5'\)-GTGYCAGCMGCCGCGGTAA-\(3'\)) and 806R
(\(5'\)-GGACTACNVGGGTWTCTAAT-\(3'\)) primers from the Earth Microbiome
Project
(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
were used to amplify the target region (Caporaso \emph{et al.}, 2012;
Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
contained on their \(5'\) end a tagged sequence. Each sample was
amplified in four parallel 25 \si{\ul} reactions of which each contained
1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7
\si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of
forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase
(New England Biolabs, USA) and 5 \si{\ng} of DNA template. Cycling
conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3
\si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45
\si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and
elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an
elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}. The four
parallel reactions volumes were pooled and PCR products were purified
using the GeneJET PCR Purification Kit (Thermo Fisher Scientific, USA)
according to the manufacturer's instructions and following the protocol
that included isopropanol addition for better small DNA fragment yield.
The column was eluted in 30 \si{\ul} of deionized water. Purified PCR
products were sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM
Laboratories, Martinsried, Germany. Before sequencing at IMGM, the
second PCR amplification of the two-step PCR procedure was performed
using primers targeting the tagged region incorporated in the first PCR.
In addition, these primers contained adapter and sample-specific index
sequences. The second PCR was carried out for 8 cycles. Beside samples,
a positive and negative control were sequenced. A negative control was
comprised of four parallel PCR reactions without DNA template, while for
a positive control a mock community composed of evenly mixed DNA
material originating from 20 bacterial strains (ATCC MSA-1002, ATCC,
USA) was used. Partial 16S rRNA sequences obtained in this study have
been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI
under accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794,
SAMEA6648809 -- SAMEA6648811.

Obtained sequences were analysed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommendations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
SILVA were exported and reclassified using mothur and the RDP
(\url{http://rdp.cme.msu.edu/}) training set (version 16) reference
files adapted for mothur (Cole \emph{et al.}, 2014). In comparison to
SILVA, RDP allows a more detailed classification of chloroplast
sequences. Based on the ATCC MSA-1002 mock community included in the
analysis a sequencing error rate of 0.009 \si{\percent} was determined,
which is in line with previously reported values for next-generation
sequencing data (Kozich \emph{et al.}, 2013; Schloss \emph{et al.},
2016). In addition, the negative control processed together with the
samples yielded only 2 sequences after sequence quality curation.

@@

Obtained sequences were analysed on the Life Science Compute Cluster
(LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
Individual sequences were assembled using MEGAHIT (version 1.1.2) (Li
\emph{et al.}, 2015) under default settings. Putative genes were
predicted from contigs longer than 200 bp using Prodigal (version 2.6.3)
(Hyatt \emph{et al.}, 2010) in metagenome mode (-p meta). Abundances of
predicted genes were expressed as Reads Per Kilobase Million (RPKM) and
calculated using the BWA algorithm (version 0.7.16a) (Li and Durbin,
2009). All predicted genes were functionally annotated using the
eggNOG-mapper (Huerta-Cepas \emph{et al.}, 2017) and eggNOG database
(version 5.0) (Huerta-Cepas \emph{et al.}, 2019). Sequence taxonomy
classification was determined using the lowest common ancestor algorithm
adapted from DIAMOND (version 0.8.36) (Buchfink \emph{et al.}, 2015) and
by searching against the NCBI non-redundant database (NR). To determine
the phylogeny, the top 10 \si{\percent} hits with an e-value \textless{}
1 × 10\textsuperscript{-5} were used (-\/-top 10). Sequence renaming,
coverage information computing and metagenomic statistics calculations
were performed using software tools from BBTools
(\url{https://jgi.doe.gov/data-and-tools/bbtools}). Metagenomic
sequences obtained in this study have been deposited in the European
Nucleotide Archive (ENA) at EMBL-EBI under accession numbers
SAMEA6648795, SAMEA6648797, SAMEA6648809 and SAMEA6648811.

@@

Isolated proteins were whole trypsin digested using the FASP
(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
according to the manufacturer's instructions with small modifications
(Wiśniewski \emph{et al.}, 2009). Prior to loading the solution onto the
column, protein pellets were solubilized in a urea sample buffer
included in the kit amended with DTT (final concentration 100
\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
remove larger particles. The first washing step after protein solution
loading was repeated twice. In addition, the centrifugation steps were
prolonged if the column was clogged. Trypsin digestion was performed on
column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
The final filtrate containing peptides was acidified with 1
\si{\percent} (final concentration) trifluoroacetic acid (TFA), freezed
at \num{-80} \si{\degreeCelsius}, lyophilized and sent to VIME -- Vienna
Metabolomics Center (University of Vienna) for metaproteomic analysis.
Peptides were re-suspended in 1 \si{\percent} (final concentration) TFA,
desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA)
according to the manufacturer's instructions and sequenced on a Q
Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher
Scientific, USA). Obtained MS/MS spectra were searched against a protein
database composed of combined sequenced metagenomes using SEQUEST-HT
engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
Fisher Scientific, USA). The target-decoy approach was used to reduce
the probability of false peptide identification. Results whose false
discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
For protein identification a minimum of two peptides and one unique
peptide were required. For protein quantification, a chromatographic
peak area-based free quantitative method was applied.

@@

Host leaves and thalli from DNA and protein isolation steps were washed
seven times in deionized water and fixed with formaldehyde (final
concentration \textasciitilde{}3 \si{\percent}). In addition,
non-treated leaves and thalli, washed three times in ASW to remove
loosely attached microbial cells, were fixed in the same concentration
of formaldehyde and used as a positive control. For-long term storage,
fixed leaves and thalli were immersed in a mixture of phosphate-buffered
saline (PBS) and ethanol (1:1) and stored at \num{-20}
\si{\degreeCelsius}. Treated and untreated segments of leaves and thalli
were stained in a 2 × solution of SYBR Green I and examined under a
Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).

@@

\leavevmode\hypertarget{ref-Li2009a}{}%
Li, H. and Durbin, R. (2009) Fast and accurate short read alignment with
Burrows--Wheeler transform. \emph{Bioinformatics} \textbf{25}:
1754--1760.

@@

\leavevmode\hypertarget{ref-Verlaque2003}{}%
Verlaque, M., Durand, C., Huisman, J.M., Boudouresque, C.-F., and Le
Parco, Y. (2003) On the identity and origin of the Mediterranean
invasive \emph{Caulerpa racemosa} (Caulerpales, Chlorophyta). \emph{Eur
J Phycol} \textbf{38}: 325--339.

@@

\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$1 \si{\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\textit{C. nodosa} and \textit{C. cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-3}
\end{figure}
